% filepath: /home/karimelbasiouni712/SWE-IV/4AL3/CareerCompass/TeamContract/TeamContract.tex
\documentclass[12pt]{article}
\usepackage[margin=0.8in]{geometry}
\usepackage{setspace}
\onehalfspacing
\usepackage{hyperref}

\begin{document}

\begin{center}
  {\Large\bfseries Team Contract}\\[10pt]
  {\normalsize Group 85}
\end{center}

\section*{1. Team Purpose / Mission}
Our team is dedicated to successfully completing the CareerCompass project by collaborating effectively, sharing knowledge, and delivering a high-quality result. The mission is to leverage each member's strengths to overcome challenges and ensure project success.

\section*{2. Roles and Responsibilities}
\begin{itemize}
  \item \textbf{Member 1:} Leads the overall coordination, communication, and ensures deadlines are met.
  \item \textbf{Member 2:} Focuses on research, development, and assists with documentation.
  \item \textbf{Member 3:} Manages testing, quality assurance, and supports integration efforts.
\end{itemize}
Each team member is responsible for proactive communication, meeting attendance, and sharing progress updates.

\section*{3. Leadership and Management}
Leadership responsibilities will be rotated on a weekly basis or as agreed upon by the team. Decisions will be discussed in team meetings and reached by a consensus, ensuring that every member’s voice is heard.

\section*{4. Conflict Resolution and Ground Rules}
\begin{itemize}
  \item If a conflict arises, the team will hold a meeting to openly discuss the issues and work towards a compromise.
  \item Decisions will be made collectively, with respect for each member’s opinion.
  \item All members agree to maintain a respectful and professional interaction at all times.
\end{itemize}

\section*{5. Additional Expectations}
The team commits to clear and open communication through regular meetings (virtual or in-person) and using our agreed-upon communication channels. Members agree to notify the team as soon as possible if any issues arise that may affect their performance.

\section*{6. Signatures}
\bigskip

\begin{tabular}{p{0.3\linewidth} p{0.3\linewidth} p{0.3\linewidth}}
\textbf{Member 1:} \newline Print Name, Sign, Date & \textbf{Member 2:} \newline Print Name, Sign, Date & \textbf{Member 3:} \newline Print Name, Sign, Date \\
\end{tabular}

\end{document}